Towards a free high quality powder X-ray diffraction dataset

### 2024-04-09 paper/00_abstract.tex
Updated abstract

We, the AiMat group at the Karlsruhe Institute of Technology (KIT), are collecting a dataset of both labeled and unlabeled experimental powder X-ray diffraction (pXRD) data, which will be made publicly available once sufficient data has been collected. We look forward to publishing a paper surveying the database and discussing its uses, and would be happy to include anyone contributing data as co-authors. You can view the current draft here: State-of-the-database

The main motivation for collecting and publishing this dataset is to assist researchers working on machine learning based approaches for pXRD analysis. Any data, even unlabeled data, can provide clues how simulations can be made to better reflect real-world experiments and labeled data can be used to test the developed algorithms. We hope to provide both types of data in a wide variety and large quantity.

While we do not collect single crystal data, we are interested in any kind of polycrystalline samples, including thin films. If your samples exhibit preferred crystallite orientation, please make mention of this in the description field of the upload form. So far, we have received upwards of 4155 unique powder X-ray diffraction patterns from institutions including:

>>>>

% Autonomous laboratory
% https://www.nature.com/articles/s41586-023-06734-w

% Accelerating material discovery
% https://www.nature.com/articles/s41524-022-00765-z

%High-throughput experiments facilitate materials innovation: A review
%https://link.springer.com/article/10.1007/s11431-018-9369-9#:~:text=High%2Dthroughput%20materials%20experiments%20including,at%20a%20fraction%20of%20cost.

%The advent of high-throughput experiments, holds the prospect of accelerating the speed materials discovery multiple times over. The synthesis and probing of novel materials is becoming increasingly efficient and automated, which has correspondingly increased the throughput of samples in these steps of the experimentation pipeline.

% You need a good initial guess, it can only optimize to the nearest minimum from the initial guess

% Comments -> Why having to do rietveld refinement sucks
% FullProf tutorial on fitting XRD peaks -> Tedious, requires lots of manual input
% https://pranabdas.github.io/research/fullprof/

% Topas tutorial -> 44 page tutorial, many boring and potentially frustrating steps, lots of points offer potential for task failure
% https://topas.webspace.durham.ac.uk/topas_user_menu/

%  Methods and Tutorials – Powder Diffraction -> Overview of rietveld refinement data
%https://neutrons.ornl.gov/sites/default/files/Methods%20and%20Tutorials%20Powder%20Diffraction.pdf

% Rietveld Refinement from Powder Diffraction Data -> 50 pg tutorial that no one wants to read, overwhelming amout of quantities/figures that needs to be manually looked at in the process
% https://www.fkf.mpg.de/4112052/cpd26.pdf

In powder x-ray diffraction, the analysis of the diffractograms still presents a significant bottleneck in high-throughput experimentation that remains to be solved, in order to match the increased throughput of the prior stages of experimentation. While Rietveld refinement can identify the phases and extract the structural information found in the diffractograms, each sample requires manual inputs and careful configuration of the analysis software from trained staff, which is infeasible for highly automated high-throughput experiments.

In the current landscape of powder x-ray diffraction analysis, machine learning has emerged as a very promising direction of research resolving this bottleneck by enabling fully automated powder diffraction analysis.
A noteable difficulty in applying machine learning to this domain is the lack of labeled datasets i.e. raw experimental diffraction patterns with accompanying structure solutions, which has relegated machine learning researches to training on simulated data. Since these simulations largely fail to accurately reflect experiment, the performance of models trained on this simulated data, also mostly fails to transfer to experiment and provide value in practice. 
We hope to remedy this by providing a freely available and easily accessible dataset of powder diffraction data, which provides machine learning researches with a large quantity of real experimental diffractograms collected from a broad range of samples. We hope that this can guide machine learning research towards more realistic simulations and potentially enable training and testing on real experimental data.

% mention below:
% uniform, easily readable

### 2024-04-09 paper/_main.tex
Updated abstract

\pdfoutput=1
\usepackage[utf8]{inputenc}

%\usepackage[
%backend=biber,
%%style=numeric,
%style=nature,
%sorting=none,
%doi=true,
%maxbibnames=99,
%date=iso,
%seconds=true,
%urldate=iso
%%citestyle=numeric-comp
%]{biblatex}

\usepackage[super,sort&compress,comma]{natbib} 
\newcommand{\supercite}[1]{\cite{#1}}

@@

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\DeclareSIUnit\angstrom{\text {Å}}

\usepackage[final]{changes}
%\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\interfootnotelinepenalty=10000
\title{Towards a high quality public powder X-ray diffraction dataset}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Pasc1al Friederich}
\author[2]{Ben Breitung}
\author[2]{Simon Schweidler}
\author[3]{Andrea Hodge}
\author[3]{Adie Alwen}
\author[4]{Mriganka Singh}
\author[4]{Carolin Sutter-Fella}
\author[4]{Tim Kodalle}
\author[4]{Maged Abdelsamie}

@@

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology, Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, 3650 McClintock Ave, Los Angeles, CA 90089, USA}
\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California, 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, 1 Cyclotron Rd., Berkeley, 94720 CA, USA}

@@

\maketitle
\begin{abstract}
    \input{00_abstract.tex}
\end{abstract}

\section{Motivation}
\input{paper/01_motivation}

\section{Existing datasets}
\input{paper/02_existing_datasets}

\section{Our dataset}
\input{paper/03_our_dataset}

\section{Contributions}
\input{paper/04_survey}

\subsection*{Acknowledgements}
\input{07_acknowledgements.tex}

>>>>

\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[super,sort\&compress,comma]{natbib}
\renewcommand{\supercite}[1]{\cite{#1}}

@@

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage[final]{changes}

\interfootnotelinepenalty=10000
\title{Towards a free high quality powder X-ray diffraction dataset}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Pasc1al Friederich}
\author[2]{Ben Breitung}
\author[2]{Simon Schweidler}
\author[3]{Andrea Hodge}
\author[3]{Adie Alwen}
\author[4,5]{Carolin Sutter-Fella}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[4]{Maged Abdelsamie}

@@

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, \protect \\3650 McClintock Ave, Los Angeles, CA 90089, USA}
\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}

@@

\maketitle
\begin{abstract}
\input{00_abstract}
\end{abstract}

\section{Motivation}\label{sec:motivation}
\input{01_motivation}

\section{Existing datasets}\label{sec:existing_datasets}
\input{02_existing_datasets}

\section{Our dataset}\label{sec:our_dataset}
\input{03_our_dataset}

\section{Contributions}\label{sec:contributions}
\input{04_contributions}
